Everthing you wanted to know about PanDA but were truly truly afraid to ask

### 2016-09-19 WMSLiterature.tex

\subsection{HTCondor Glidein and GlideinWMS}
The HTCondor Glidein system  as part of
the HTCondor [153] software ecosystem. The HTCondor
Glidein is a pilot-based system to aggregate DCRs with
heterogeneous middleware into HTCondor resource pools.
The logical components of HTCondor relevant to the
Glidein system are: 
\begin{itemize}
\item \emph{Schedd}, a daemon that implement a queuing system that holds workload tasks;
\item \emph{Startd}, a  daemon to control the DCR resources. 
\item \emph{Collector}, holds references to all the active
Schedd/Startd daemons; 
\item \emph{Negotiator} matches tasks queued in a Schedd to resources handled by a Startd.
\end{itemize}
In
HTCondor Glidein has been complemented by Glidein-WMS, a Glidein-based workload management system thatautomates deployment and management of Glideins on multiple types of DCR middleware. GlideinWMS builds upon the
HTCondor Glidein system by adding the following logical components: 
\begin{itemize}
\item \emph{Glidein Factories} that submit tasks to the DCRs middleware;
\item a set of \emph{Virtual Organizations (VO) Frontend} daemons that match the tasks on one or more Schedd to the resource attributes;
\item a \emph{Collector} that holds references to all the active Glidein Factories and VO Frontend daemons. 
\end{itemize}

@@

Both HTCondor Glidein and GlideWMS rely on one or more HTCondor Collectors to match task requirements and resource properties, represented as ClassAds. This
matching can be evaluated right before the scheduling of the task. In this way, late binding is achieved but early binding remains unsupported.

>>>>

\subsection{HTCondor Glidein and GlideinWMS}
The HTCondor Glidein system  as part of the HTCondor software ecosystem. The HTCondor Glidein is a pilot based system to aggregate DCRs with heterogeneous middleware into HTCondor resource pools.
Condor is based on daemons collaborating by exchanging messages over the network. We can isolate four main logical components:
\begin{itemize}
\item \emph{Schedd}, implements a queuing system that holds workload tasks;
\item \emph{Startd}, controls the DCR resources. 
\item \emph{Collector}, holds references to all the active
Schedd/Startd daemons; 
\item \emph{Negotiator} matches tasks queued in a Schedd to resources handled by a Startd.
\end{itemize}
Glidein-WMS has been developed to integrate HTCondor Glidein to  automate the deployment and management of Glideins on multiple types of DCR middleware. 
The integration required three additional logical components: 
\begin{itemize}
\item \emph{Glidein Factories} that submit tasks to the DCRs middleware;
\item a set of \emph{Virtual Organizations (VO) Frontend} daemons that match the tasks on one or more Schedd to the resource attributes;
\item a \emph{Collector} that holds references to all the active Glidein Factories and VO Frontend daemons. 
\end{itemize}

@@

Both HTCondor Glidein and GlideWMS rely on one or more HTCondor Collectors to match task requirements and resource properties, represented as ClassAds. This matching can be evaluated right before the scheduling of the task. In this way, late binding is achieved but early binding remains unsupported.
\begin{table*}
\begin{center}
\begin{tabular}{llllllll}
  \hline
Pilot System  &Logical Components& Execution Strategy & Binding  & Workload Definition  &  Broker  & \\
\hline
Alien & Central services, site services & pull, pilot via Co-pilot & Late & JDL & Condor (ClassAd attributes) &\\
DIRAC & Services, Agents& pull, pilot-based & Late & JDL, WF (TMS) & Condor(ClassAd attributes) &\\
glidein WMS& Daemons & pull, pilot-based  & Late & Pegaus, DAGMan & Condor(ClassAd attributes) &\\
\hline
\end{tabular}
\end{center}
\caption{Comparison of the three Workload and Data Management Systems}\label{tab:Summary}
\end{table*}

### 2016-09-19 paper.tex

% ----------------------------------------------------------------------------
% I
% ----------------------------------------------------------------------------
\section{Introduction}\label{sec:intro}
The Large Hadron Collider (LHC) is created  to explore the fundamental  properties  of matter for the next decades. Since LHC start-up in 2009, the experiments has collected and distributed hundreds  of  petabytes  of  data worldwide to hundreds of computer centers. Thousands of physicists analyze peta-scale  data volumes daily. One of  the LHC experiments  the ATLAS  [1]  utilizes the PanDA  workload management system [4] (WMS) for distributed data processing and analysis. The ATLAS Computing model [2] is based on a Grid paradigm [3], with multilevel, hierarchically distributed computing  and storage resources. Acronym  PanDa stands for Production  and Distributed Analysis, the system  has  been developed to meet growing ATLAS production and analysis requirements for a data-driven  workload  management system capable of operating at LHC data processing scale.
PanDA has a highly scalable architecture. Scalability  has been demonstrated in ATLAS through the rapid increase in usage over the past several years of operations, and is expected to meet the continuously growing number of jobs over the next decade. PanDA  was designed to have the flexibility to adapt to emerging computing  technologies in processing,  storage, networking  as well as the underlying software stack (middle- ware). This flexibility has also been successfully demonstrated through the past six years of evolving  technologies adapted by computing centers in ATLAS which span many continents and yet are seamlessly integrated into PanDA.
PanDA  manages a wide spectrum of workloads, ranging from raw data processing to Monte Carlo simulation and user anal- ysis, while actively evolving to meet rapidly changing science needs.  Today, it  serves several  thousand users, successfully managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU cores  which process  more than a million jobs per day, Figure 1 shows  a  summary   of daily completed jobs on ATLAS Grid for the past 12 month.
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figure/DailyJobs.png}
\caption{Daily completed jobs on ATLAS Grid\label{fig:daily}}
\end{center}
\end{figure}

@@

\begin{enumerate}
  \item Brief overview of distributed CI.
  \item Workload vs Workflow
  \item Distinguish between WFMS and WLMS!!!!
  \begin{enumerate}
    \item Role of Workload MS: help “hide” and federate heterogeneous resources.
  \end{enumerate}
\end{enumerate}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figure/PandaArch.jpg}
\caption{Schematic view of the PanDA system\label{fig:architecture}}
\end{center}
\end{figure}
\subsection{PanDA Server}
The PanDA server is the main component of the system, It provides  a task queue managing all job information centrally. The PanDA  server receives jobs through the client interface into the task queue, upon which a brokerage module  operates to prioritize  and assign work on the basis of job type, priority, input data and its locality, and available CPU resources. The PanDA  server operates as a web service. It runs on Apache web server, interacting with the back-end database running  on separate servers. It uses the Apache worker  model, with many independent   processes  handling client requests  in  parallel. Since all  state is maintained  in  the central database,  the PanDA Server application instance itself is stateless. Currently production version of PanDA utilizes Oracle database backend, but PanDA  can also work with MySQL family of databases.

@@

\subsection{Use of Supercomputers with PanDA}
The idea is to extend the ATLAS Computing Model beyond the Grid into the domains of supercomputers. Modern High Performance Computing  (HPC) involves  a very large number of worker nodes  connected  through a  high-speed   network. The worker nodes can have  multicore CPUs that can be augmented with massively parallel Graphics Processing Units (GPUs) or other types of coprocessors. On an HPC machine, a  typical job is highly parallel and connected,  where each core calculates  a small part of the problem and coordinates the activity via message passing interface (MPI). While this is different from the HEP computing paradigm, where jobs are independent, it still shares common  features  such as the use of parallelization. It is not a requirement  that HPC machines are able to run any possible task, nor is it relevant how many kinds of job types that can be run. What matters is the total number of cycles that can be offloaded from the Grid. Standard  ATLAS  workflow  can not be easily ported to supercomputers   due  to   several   complications such  as specialized worker  node  setups, no  outbound network connections,  limited memory per node, custom operating systems,  etc. A reorganization  of the standard  workflow is therefore needed.
The most suitable task for HPCs in HEP is event generation. Event generators are mostly computational,  stand-alone code, with few input requirements.  There is no need  to stage-in much data. The event generation  in ATLAS corresponds  to
10-15\% of all jobs on the Grid. A more difficult workload to adapt  for HPCs is detector  simulation. Simulation tasks tend to be tied to the framework of a particular simulation and often require  database access for geometry and relevant run conditions.  However,  there is a  lot to be gained  since simulation  tasks correspond to 2/3 of the Grid capacity.
\subsection{Interfacing PanDA with Titan}
The Titan supercomputer, current number two (number one from November 2012 until June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing  Facility (OLCF) in Oak Ridge National Laboratory, US. Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
Integration with Titan is the current focus for PanDA develop- ers. The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface [6]. This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world. The pilot can communicate with the PanDA server from the front-end machine. The interactive front-end  machines and the worker nodes  use  a shared  file system  which makes  it  easy  for the pilot to stage-in  any input files that are required by the payload and stage-out the produced output files after the payload has finished at the end of the job. ATLAS Tier 1 computing  center at Brookhaven National Lab is currently  used for data stage out from Titan but in principle  that can be any Grid site.
1) HPC Backfill: HPC facilities are geared towards  large scale jobs by design. Time allocation on an HPC is competitive and large projects are often preferred. About 10% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources. The worker nodes
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figure/PandaInterfaceWithHPC.png}
\caption{Schematic view of the PanDA system\label{fig:architecture}}
\end{center}
\end{figure}

>>>>

% ----------------------------------------------------------------------------
% I
% ----------------------------------------------------------------------------
\section{Introduction}\label{sec:intro}
The Large Hadron Collider (LHC) is created  to explore the fundamental  properties  of matter for the next decades. Since LHC start-up in 2009, the experiments has collected and distributed hundreds  of  petabytes  of  data worldwide to hundreds of computer centers. Thousands of physicists analyze peta-scale  data volumes daily. One of  the LHC experiments  the ATLAS  [1]  utilizes the PanDA  workload management system [4] (WMS) for distributed data processing and analysis. The ATLAS Computing model [2] is based on a Grid paradigm [3], with multilevel, hierarchically distributed computing  and storage resources. Acronym  PanDa stands for Production  and Distributed Analysis, the system  has  been developed to meet growing ATLAS production and analysis requirements for a data-driven  workload  management system capable of operating at LHC data processing scale.
PanDA has a highly scalable architecture. Scalability  has been demonstrated in ATLAS through the rapid increase in usage over the past several years of operations, and is expected to meet the continuously growing number of jobs over the next decade. PanDA  was designed to have the flexibility to adapt to emerging computing  technologies in processing,  storage, networking  as well as the underlying software stack (middle- ware). This flexibility has also been successfully demonstrated through the past six years of evolving  technologies adapted by computing centers in ATLAS which span many continents and yet are seamlessly integrated into PanDA.
PanDA  manages a wide spectrum of workloads, ranging from raw data processing to Monte Carlo simulation and user anal- ysis, while actively evolving to meet rapidly changing science needs.  Today, it  serves several  thousand users, successfully managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU cores  which process  more than a million jobs per day, Figure 1 shows  a  summary   of daily completed jobs on ATLAS Grid for the past 12 month.
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/DailyJobs.png}
\caption{Daily completed jobs on ATLAS Grid\label{fig:daily}}
\end{center}
\end{figure}

@@

\begin{enumerate}
  \item Brief overview of distributed CI.
  \item Workload vs Workflow. (DONE)
  \item Distinguish between WFMS and WLMS.
  \begin{enumerate}
    \item Role of Workload MS: help ``hide'' and federate heterogeneous resources.
  \end{enumerate}
\end{enumerate}

\paragraph*{Workload Vs Workflow} The term ``workflow'' is used in many
disciplines with different meaning. In the field of scientific computing,
``workflow'' assumes different meanings depending on the characteristics of
the computation, of the software tools used to support this computation, and
of the resources on which it is performed. Further, a workflow may indicate a
whole application, a description of the computational process of that
application or, more commonly, a series of tasks related by data
dependences.

The lack of a consistent and shared definition of ``workflow'' hinders the
understanding of its properties and its relations with related concepts. For
example, we need to clarify the difference between ``workflow'', ``workload'',
``task'', or ``job'' but also between workflow template and instance, or
data-flow and control-flow. This clarification is precondition to specify
properly the design of software systems that support the execution of
scientific applications.

In this paper, a workflow is a set of tasks alongside a description of their
properties and relations. Each task may have an arbitrary number of properties
like number of cores, executables, or input/output files. Tasks may have
precedence relations depending on their data dependences or any other type of
dependency mandated by the application algorithms. Tasks with precedence
relations have to be executed serially, otherwise tasks can be executed
concurrently. A workload is defined as a set of tasks that can be executed
concurrently. As such, a workflow can be composed by a set of workloads.

The terms ``task'' and ``job'' are also used inconsistently across communities
that perform scientific computing. In this paper, a task is a \ldots
\mtnote{Are we going to use task and job as defined in Jedi and Panda Server?}

Usually, workflows are represented as graphs where tasks are vertices and
their relations edges~\cite{}. Often, graphs are supposed to be acyclic but
graphs with cycles have been used to represent workflows of workflows. In this
paper, a workflow template is a type of graph while a workflow instance is a
type of graph with a specific set of vertices and edges. Further, a graph is
``abstract'' when no resource properties are available for all vertices,
``concrete'' otherwise~\cite{}.

The distinction between an abstract and concrete graph highlights an important
element of the execution process of a workflow. Workflows can be specified
without indicating on what resources they should be executed. Accordingly, an
abstract workflow can be interpreted as a formal description of the
requirements for the workflow execution. Once these requirements are matched
by a set of resource capabilities, an abstract workflow becomes a concrete
workflow, i.e. a formal description of the workflow execution process and its
execution environment.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/PandaArch.jpg}
\caption{Schematic view of the PanDA system\label{fig:architecture}}
\end{center}
\end{figure}
\subsection{PanDA Server}
The PanDA server is the main component of the system, It provides  a task queue managing all job information centrally. The PanDA  server receives jobs through the client interface into the task queue, upon which a brokerage module  operates to prioritize  and assign work on the basis of job type, priority, input data and its locality, and available CPU resources. The PanDA  server operates as a web service. It runs on Apache web server, interacting with the back-end database running  on separate servers. It uses the Apache worker  model, with many independent   processes  handling client requests  in  parallel. Since all  state is maintained  in  the central database,  the PanDA Server application instance itself is stateless. Currently production version of PanDA utilizes Oracle database backend, but PanDA  can also work with MySQL family of databases.

@@

\subsection{Use of Supercomputers with PanDA}
The idea is to extend the ATLAS Computing Model beyond the Grid into the domains of supercomputers. Modern High Performance Computing  (HPC) involves  a very large number of worker nodes  connected  through a  high-speed   network. The worker nodes can have  multicore CPUs that can be augmented with massively parallel Graphics Processing Units (GPUs) or other types of coprocessors. On an HPC machine, a  typical job is highly parallel and connected,  where each core calculates  a small part of the problem and coordinates the activity via message passing interface (MPI). While this is different from the HEP computing paradigm, where jobs are independent, it still shares common  features  such as the use of parallelization. It is not a requirement  that HPC machines are able to run any possible task, nor is it relevant how many kinds of job types that can be run. What matters is the total number of cycles that can be offloaded from the Grid. Standard  ATLAS  workflow  can not be easily ported to supercomputers   due  to   several   complications such  as specialized worker  node  setups, no  outbound network connections,  limited memory per node, custom operating systems,  etc. A reorganization  of the standard  workflow is therefore needed.
The most suitable task for HPCs in HEP is event generation. Event generators are mostly computational,  stand-alone code, with few input requirements.  There is no need  to stage-in much data. The event generation  in ATLAS corresponds  to
10-15\% of all jobs on the Grid. A more difficult workload to adapt  for HPCs is detector  simulation. Simulation tasks tend to be tied to the framework of a particular simulation and often require  database access for geometry and relevant run conditions.  However,  there is a  lot to be gained  since simulation  tasks correspond to 2/3 of the Grid capacity.
\subsection{Interfacing PanDA with Titan}
The Titan supercomputer, current number two (number one from November 2012 until June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing  Facility (OLCF) in Oak Ridge National Laboratory, US. Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
Integration with Titan is the current focus for PanDA develop- ers. The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface [6]. This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world. The pilot can communicate with the PanDA server from the front-end machine. The interactive front-end  machines and the worker nodes  use  a shared  file system  which makes  it  easy  for the pilot to stage-in  any input files that are required by the payload and stage-out the produced output files after the payload has finished at the end of the job. ATLAS Tier 1 computing  center at Brookhaven National Lab is currently  used for data stage out from Titan but in principle  that can be any Grid site.
1) HPC Backfill: HPC facilities are geared towards  large scale jobs by design. Time allocation on an HPC is competitive and large projects are often preferred. About 10% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources. The worker nodes
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/PandaInterfaceWithHPC.png}
\caption{Schematic view of the PanDA system\label{fig:architecture}}
\end{center}
\end{figure}
